Discrete Quantum Theories and Computing

### 2018-11-20 dissertation.tex
dissertation.lyx "2.2 Quantum Probability" Updating...

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Given a finite sample space~$\Omega$ representing all possible outcomes,
and its power set $2^{\Omega}$ as the classical event space, a classical
probability measure~$\mu$ maps an event~$E$ to a number $\mu\left(E\right)\in\left[0,1\right]$
specifying how likely one of the outcome in $E$ actually happens.
To maintain the coherence, we want $\mu$ subject to the following
constraints. When $E$ contains no outcome, $E$ is the least likely
to happen, i.e., $\mu\left(\emptyset\right)=0$. When $E$ contains
all outcomes, $E$ is the most likely to happen, i.e., $\mu\left(\Omega\right)=1$.
If we know the likelihood of an event, $\mu\left(E\right)$, it is
natural to assume the likelihood of the event’s complement is just
the opposite of $\mu\left(E\right)$, i.e., $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$.
For each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$,
the likelihood of the union event is additive, $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$.

The previous abstraction

@@

We now explicitly define $\varphi$. First, $\events'$ is called
a \emph{subspace} of the set of events~$\events$ if $\events'$
contains the projectors $\mathbb{0}$ and $\mathbb{1}$ and is closed
under complements, sums, and products. In particular, for any projector
$P\in\events'$, we have $\mathbb{1}-P\in\events'$ and for each pair
of commuting projectors $P_{0}\in\events'$ and $P_{1}\in\events'$,
we have $P_{0}P_{1}$ and $P_{0}+P_{1}-P_{0}P_{1}\in\events'$. Given
a mutually commuting subspace~$\eventsC$ whose elements are diagonalizable
by a common orthonormal basis $\Omega=\left\{ \ket{0},\ket{1},\ldots,\ket{D-1}\right\} $,
we define the function $\varphi:2^{\Omega}\rightarrow\events$ maps
any set~$E$ to the sum of the projectors formed by elements in $E$,
i.e., 
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\,.\label{eq:pullback-function}
\end{equation}
If we follow the convention $\varphi\left(\emptyset\right)=\mathbb{0}$,
this function $\varphi$ naturally sends the set operations to the
corresponding projector operations: $\varphi\left(\Omega\right)=\mathbb{1}$,
$\varphi\left(\overline{E}\right)=\mathbb{1}-\varphi\left(E\right)$,
and
\begin{align}
\varphi\left(E_{0}\cap E_{1}\right) & =\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,, & \varphi\left(E_{0}\cup E_{1}\right) & =\varphi\left(E_{0}\right)+\varphi\left(E_{1}\right)-\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\label{eq:pullback-properties}
\end{align}
for each pair of set $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.
Therefore, given a QIVPM~$\bar{\mu}$, the function $\varphi^{*}\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
defined by precomposition $\left(\varphi^{*}\bar{\mu}\right)\left(E\right)=\bar{\mu}\left(\varphi\left(E\right)\right)$
is the pullback of $\bar{\mu}$ by $\varphi$, and is a classical
IVPM naturally. Moreover, if $\mu:\events\rightarrow\left[0,1\right]$
is a quantum probability measure satisfying Eqs.~(\ref{eq:QuantumProbability})
and (\ref{eq:QuantumProbability-Addition}), the pullback of $\mu$,
$\varphi^{*}\mu:2^{\Omega}\rightarrow\left[0,1\right]$, can also
be defined by precomposition $\left(\varphi^{*}\mu\right)\left(E\right)=\mu\left(\varphi\left(E\right)\right)$,
and is a classical probability measure as well.

>>>>

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps an event~$E$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcome in $E$ actually happens. To maintain the
coherence, $\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
and $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

Since the previous abstraction doesn't specify how to generate the
outcomes, the generating process could well be a quantum experiment.
Let us prepare a beam of one kind of spin~$1$ particles with basis
vectors $\ket{0}$, $\ket{1}$, and $\ket{2}$. In principle, this
beam can be split by a Stern-Gerlach type experiment according to
the eigenvalues of the observable $\mathbf{O}_{0}=0\,\proj{0}+1\,\proj{1}+2\,\proj{2}$,
and the states corresponding to the split beam are $\ket{0}$, $\ket{1}$,
and $\ket{2}$ \cite{:/content/aip/journal/jmp/21/1/10.1063/1.524312,peres1995quantum}.
Instead of a beam, if only one particle is sent to the beam splitter,
the state after the experiment is one of $\ket{0}$, $\ket{1}$, and
$\ket{2}$, and each state has a probability. In the language of our
previous abstraction, the set of outcomes is $\Omega_{0}=\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
All possible events are $\emptyset$, $\left\{ \ket{0}\right\} $,
$\left\{ \ket{1}\right\} $, $\left\{ \ket{2}\right\} $, $\left\{ \ket{0},\ket{1}\right\} $,
$\left\{ \ket{0},\ket{2}\right\} $, $\left\{ \ket{1},\ket{2}\right\} $,
and $\Omega_{0}$. This experiment defines a classical probability
measure $\mu_{0}:2^{\Omega_{0}}\rightarrow\left[0,1\right]$.

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
similar to the previous one. While this experiment send exactly the
same particle, we use a beam splitter corresponding to the observable
$\mathbf{O}_{1}=0\,\proj{\ps}+1\,\proj{\ms}+2\,\proj{2}$, where $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$
and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$. The outcome of
this experiment is $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $,
and it defines another probability measure $\mu_{1}:2^{\Omega_{1}}\rightarrow\left[0,1\right]$.
Although $\mu_{0}$ and $\mu_{1}$ are different probability measures,
the probability of the same event are experimentally found to be the
same, i.e., $\mu_{0}\left(\emptyset\right)=\mu_{1}\left(\emptyset\right)$
and $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$.
In general, when the particle and the event are the same, so does
their probability no matter how we set up the beam splitters.

Since the probability induced by the different beam splitters are
correlated, it is more natural to define the set of quantum event~$\events$
as the union of all possible event spaces using different beam splitters.
Since we are focus on the probability instead of outcome, and the
probabilities of events $\left\{ \ket{0},\ket{1}\right\} $ and $\left\{ \ket{\ps},\ket{\ms}\right\} $
are always the same, it is reasonable to identify them as the same
event. To simplify the notation, given any classical event $E\subseteq\Omega$,
its corresponding quantum event is defined by 
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\,.\label{eq:pullback-function}
\end{equation}
Following this definition, it is easy to check that not only 
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)
\end{equation}
but every classical event mapping the same projector must also have
the same probability. If we follow the convention $\varphi\left(\emptyset\right)=\mathbb{0}$,
this function $\varphi$ naturally sends the set operations to the
corresponding projector operations: $\varphi\left(\Omega\right)=\mathbb{1}$,
$\varphi\left(\overline{E}\right)=\mathbb{1}-\varphi\left(E\right)$.
For each pair of set $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$,
\begin{align}
\varphi\left(E_{0}\cap E_{1}\right) & =\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,, & \varphi\left(E_{0}\cup E_{1}\right) & =\varphi\left(E_{0}\right)+\varphi\left(E_{1}\right)-\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,.\label{eq:pullback-properties}
\end{align}
If $E_{0}$ and $E_{1}$ are disjoint, then $\varphi\left(E_{0}\right)$
and $\varphi\left(E_{1}\right)$ are orthogonal, where two projectors
$P_{0}$ and $P_{1}$ are called \emph{orthogonal} if $P_{0}P_{1}=\mathbb{0}$.

, and consider the quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
as a function assigning the likelihood to every quantum event

@@

We now explicitly define $\varphi$. First, $\events'$ is called
a \emph{subspace} of the set of events~$\events$ if $\events'$
contains the projectors $\mathbb{0}$ and $\mathbb{1}$ and is closed
under complements, sums, and products. In particular, for any projector
$P\in\events'$, we have $\mathbb{1}-P\in\events'$ and for each pair
of commuting projectors $P_{0}\in\events'$ and $P_{1}\in\events'$,
we have $P_{0}P_{1}$ and $P_{0}+P_{1}-P_{0}P_{1}\in\events'$. Given
a mutually commuting subspace~$\eventsC$ whose elements are diagonalizable
by a common orthonormal basis $\Omega=\left\{ \ket{0},\ket{1},\ldots,\ket{D-1}\right\} $,
we define the function $\varphi:2^{\Omega}\rightarrow\events$ maps
any set~$E$ to the sum of the projectors formed by elements in $E$
as in Eq.~(\ref{eq:pullback-function}). \todo{Review this part
because Eq.~(\ref{eq:pullback-function}) is moved to Sec.~\ref{sec:fuzzy}.
}Therefore, given a QIVPM~$\bar{\mu}$, the function $\varphi^{*}\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
defined by precomposition $\left(\varphi^{*}\bar{\mu}\right)\left(E\right)=\bar{\mu}\left(\varphi\left(E\right)\right)$
is the pullback of $\bar{\mu}$ by $\varphi$, and is a classical
IVPM naturally. Moreover, if $\mu:\events\rightarrow\left[0,1\right]$
is a quantum probability measure satisfying Eqs.~(\ref{eq:QuantumProbability})
and (\ref{eq:QuantumProbability-Addition}), the pullback of $\mu$,
$\varphi^{*}\mu:2^{\Omega}\rightarrow\left[0,1\right]$, can also
be defined by precomposition $\left(\varphi^{*}\mu\right)\left(E\right)=\mu\left(\varphi\left(E\right)\right)$,
and is a classical probability measure as well.
